Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-03-28 methods/subsections/machine learning.tex
Updates from Overleaf

\chapter{Methods}

We propose using a triple siamese network to train a model for music similarity retrieval. Specifically, we aim to minimize the loss function between the anchor, positive, and negative samples by using online triplet mining.

The positive samples are created by transforming and augmenting the original samples while preserving their perceptual underlying musical language. This is achieved using techniques such as time-stretching, pitch-shifting, and adding noise to the samples. By doing so, we aim to enhance the diversity of the positive samples and improve the model's ability to retrieve similar-sounding samples that share the same underlying musical language.

On the other hand, the negative samples are created by shuffling and scrambling the original samples while maintaining their production texture. This is achieved using random segment selection, time-domain filtering, and frequency-domain masking. The goal is to maintain the production texture of the samples while disrupting their underlying composition.

Using the proposed triple siamese network and online triplet mining, we aim to train a model that can effectively distinguish between similar and dissimilar samples while considering their underlying composition and production texture.

\section{Materials}

@@

\subsection{Neural Network (NN)}

%Single
\input{figures/cnn/Single Neuron}
%NN
\input{figures/cnn/Network_graph}

%Backpropagation
\input{figures/cnn/Multi-layer_NN}

%CNN
A high-level illustration of a simple convolutional neural network indicating convolutional layers, pooling layers, and fully-connected layers without details on the number of channels or neurons per layer or the input data size can be seen in XXX.
\input{figures/cnn/CNN_diagram}

%Single CNN
\input{figures/cnn/Single_CNN}

%Single pooling layer
\input{figures/cnn/Single_Pooling_Layer}

%Backpropagation
\input{figures/cnn/Back_propagation}

%My model

\input{figures/my model/my model}

>>>>

Machine Learning (ML), particularly Deep Learning, has significantly advanced the Music Information Retrieval (MIR) field by enabling automatic feature learning, improving classification and analysis tasks, and facilitating novel applications like music generation and source separation. Adopting techniques like Convolutional Neural Networks (CNNs), Recurrent Neural Networks (RNNs), and generative models has led to better performance in tasks such as music classification, tagging, recommendation, transcription, synthesis, and analysis. As a result, ML has unlocked the potential of MIR, revolutionizing how we interact with, understand, and create music.

\section{Artificial Neural Networks (ANNs)}
A neural network, formally known as an artificial neural network (ANN), is a computational model inspired by the structure and functionality of biological neural networks. It is a mathematical framework designed to process information and perform tasks like pattern recognition, classification, and regression by mimicking how biological neurons and synapses operate. ANNs comprise interconnected layers of artificial neurons or nodes, each performing a mathematical transformation on the input data to generate an output. The connections between these neurons, often called "weights," are adjusted during learning to optimize the network's performance on a given task.

\subsection{About ANNs}
The concept of neural networks can be traced back to the early 20th century, with the pioneering work of Warren McCulloch and Walter Pitts in 1943. They proposed a mathematical model of a neuron known as the McCulloch-Pitts neuron, which provided a foundation for future research in artificial intelligence. However, Frank Rosenblatt's invention of the perceptron in 1957 marked the beginning of modern neural networks. The perceptron is a single-layer neural network capable of learning linearly separable patterns.

In a more technical description, an ANN consists of an input layer, one or more hidden layers, and an output layer. Each neuron in a layer receives input from the previous layer and computes an output value using an activation function. Some standard activation functions include the sigmoid, hyperbolic tangent (tanh), and rectified linear unit (ReLU). The output of the last layer represents the prediction or classification result.

%Single neuron
\input{figures/neural networks/Single Neuron}
%NN
\input{figures/neural networks/Network_graph}
%Multi-layer_NN
\input{figures/neural networks/Multi-layer_NN}
%Backpropagation
\input{figures/neural networks/Back_propagation}

@@

%CNN
A high-level illustration of a simple convolutional neural network indicating convolutional layers, pooling layers, and fully-connected layers without details on the number of channels or neurons per layer or the input data size can be seen in XXX.
\input{figures/neural networks/CNN_diagram}

%Single CNN
\input{figures/neural networks/Single_CNN}

%Single pooling layer
\input{figures/neural networks/Single_Pooling_Layer}

### 2023-03-28 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

SSL, or self-supervised learning, is a technique to learn latent representations from large-scale data without relying on human annotations. Instead, it solves pretext tasks by creating different views of an object that are highly correlated. An SSL model is then trained to generalize objects' representations in a latent high-dimensional space. To achieve this, the model contrasts the representations of the same object with other objects (defined as negative samples) during training. By doing so, a contrastive SSL model is expected to produce more distinctive representations.

TRIPLET LOSS

>>>>

\section{Model architecture and training strategy}
Self-Supervised Learning (SSL) is a technique to learn latent representations from large-scale data without relying on human annotations. Instead, it solves pretext tasks by creating different views of an object that are highly correlated. An SSL model is then trained to generalize objects' representations in a latent high-dimensional space. To achieve this, the model contrasts the representations of the same object with other objects (defined as negative samples) during training. By doing so, a contrastive SSL model is expected to produce more distinctive representations.

\section{}

We propose using a Triple Siamese Network to train a model for music similarity retrieval. Specifically, we aim to minimize the loss function between the anchor, positive, and negative samples by using online triplet mining.

The positive samples are created by transforming and augmenting the original samples while preserving their perceptual underlying musical language. This is achieved using techniques such as time-stretching, pitch-shifting, and adding noise to the samples. By doing so, we aim to enhance the diversity of the positive samples and improve the model's ability to retrieve similar-sounding samples that share the same underlying musical language.

On the other hand, the negative samples are created by shuffling and scrambling the original samples while maintaining their production texture. This is achieved using random segment selection, time-domain filtering, and frequency-domain masking. The goal is to maintain the production texture of the samples while disrupting their underlying composition.

Using the proposed triple siamese network and online triplet mining, we aim to train a model that can effectively distinguish between similar and dissimilar samples while considering their underlying composition and production texture.
